Titeln på avhandlingens språk

### 2022-12-05 appendices.tex
jaya

We first borrow the idea of checkpointing density from \textcite{Satoshi1992Optimal} and define an instantaneous sampling rate function $r(t)$ which is related to $\{t_n\}$ such that,
\begin{alignat}{1}
{\int_{t_{n-1}}^{t_n}}r(t)\dif t=1,\;\forall n\geq1\label{rt}
\end{alignat}
Note that, for periodic sampling, this function is a constant, equal to the sampling frequency.
In the aperiodic case, we find $r^*(t)$, the $r(t)$ that minimizes $\mathcal{E}$.
By construction, the number of samples taken up to any time instant can be computed directly by computing the area under $r(t)$.
Thus, we obtain the expected number of samples $\mathbb{E}[\mathcal{S}]$ as
\begin{alignat}{1}
\mathbb{E}[\mathcal{S}]&=\int_{t=0}^{\infty}\bigg(\!\int_{x=0}^{t}\!\!\!\!r(x)\,\mathrm{d}x\bigg)f_{\mathcal{T}}(t)\,\mathrm{d}t.\label{Es}
\end{alignat}

@@

As per the Euler-Lagrange equation from the calculus of variations~\cite{Bellman1954Dynamic,Arfken2015Calculus}, the extreme value of $\mathcal{E}$ is obtained at 
\begin{alignat}{1}
r^*(t)&=\sqrt{\dfrac{\beta f_\mathcal{T}(t)}{2\alpha\bar{F}_\mathcal{T}(t)}}.\\
\intertext{Thus, for a Rayleigh distributed $\mathcal{T}$ with parameter $\sigma$,}
r^*(t)&=\sqrt{\dfrac{\beta t}{2\alpha\sigma^2}}\\
\Rightarrow &\int_{t_n}^{t_{n+1}}\!\!\!\sqrt{\dfrac{\beta t}{2\alpha\sigma^2}}\,\mathrm{d}t=1,\;\forall n\geq1\nonumber\tag{from \eqref{rt}}\\
\Rightarrow &\;t_{n+1}^{\frac{3}{2}}-t_{n}^{\frac{3}{2}}=3\sigma\!\sqrt{\tfrac{\alpha}{2\beta}}\nonumber\\
\intertext{We have $t_0=0$. Substituting $n=1,2,\dots$, in order, in the above equation provides us our final result}
&\;t_n=\Big(3\sigma\!\sqrt{\tfrac{\alpha}{2\beta}}\Big)^{\frac{2}{3}}n^{\frac{2}{3}}\label{tnRayleigh}.
x\end{alignat}
% Note that throughout the paper we use a exGaussian distribution but not Rayleigh. This is because, 
Note that, due to the close similarity in their density functions, the task times can be approximated fairly equally to an exponentially modified Gaussian distribution as well as a Rayleigh distribution, with the former attracting more attention from works like \cite{Rohrer1994analysis,Palmer2011shapes,Marmolejo2022generalised}. This is the reason why the above results are applicable in this work where we have predominantly considered exGaussian distribution. Furthermore, we have also verified the closeness of the results as well as the validity of the approximations made in the proofs using distribution fitting and simulations.

>>>>

We first borrow the idea of checkpointing density from~\cite{Satoshi1992Optimal} and define an instantaneous sampling rate function $r(t)$ which is related to $\{t_n\}$ such that,
\begin{alignat}{1}
{\int_{t_{n-1}}^{t_n}}r(t)\dif t=1,\;\forall n\geq1\label{rt}
\end{alignat}
Note that, for periodic sampling, this function is a constant, equal to the sampling frequency.
In the aperiodic case, we find $r^*(t)$, the $r(t)$ that minimizes $\mathcal{E}$.
By construction, the number of samples taken up to any time instant can be computed directly by computing the area under $r(t)$.
Thus, we obtain the expected number of samples $\mathbb{E}[\mathcal{S}]$ as
\begin{alignat}{1}
\mathbb{E}[\mathcal{S}]&=\int_{t=0}^{\infty}\bigg(\!\int_{x=0}^{t}\!\!\!\!r(x)\,\mathrm{d}x\bigg)f_{\mathcal{T}}(t)\,\mathrm{d}t.\label{Es}
\end{alignat}

@@

As per the Euler-Lagrange equation from the calculus of variations~\cite{Bellman1954Dynamic,Arfken2015Calculus}, the extreme value of $\mathcal{E}$ is obtained at 
\begin{alignat}{1}
r^*(t)&=\sqrt{\dfrac{\beta f_\mathcal{T}(t)}{2\alpha\bar{F}_\mathcal{T}(t)}}.\\
\intertext{Thus, for a Rayleigh distributed $\mathcal{T}$ with parameter $\sigma$,}
r^*(t)&=\sqrt{\dfrac{\beta t}{2\alpha\sigma^2}}\\
\Rightarrow &\int_{t_n}^{t_{n+1}}\!\!\!\sqrt{\dfrac{\beta t}{2\alpha\sigma^2}}\,\mathrm{d}t=1,\;\forall n\geq1\nonumber\tag{from \eqref{rt}}\\
\Rightarrow &\;t_{n+1}^{\frac{3}{2}}-t_{n}^{\frac{3}{2}}=3\sigma\!\sqrt{\tfrac{\alpha}{2\beta}}\nonumber\\
\intertext{We have $t_0=0$. Substituting $n=1,2,\dots$, in order, in the above equation provides us our final result}
&\;t_n=\Big(3\sigma\!\sqrt{\tfrac{\alpha}{2\beta}}\Big)^{\frac{2}{3}}n^{\frac{2}{3}}\label{tnRayleigh}.
x\end{alignat}
% Note that throughout the paper we use a exGaussian distribution but not Rayleigh. This is because, 
Note that, due to the close similarity in their density functions, the task times can be approximated fairly equally to an \acl{exGaussian} distribution as well as a Rayleigh distribution, with the former attracting more attention from works like~\cite{Rohrer1994analysis,Palmer2011shapes,Marmolejo2022generalised}.
This is the reason why the above results are applicable in this work where we have predominantly considered \ac{exGaussian} distribution.
Furthermore, we have also verified the closeness of the results as well as the validity of the approximations made in the proofs using distribution fitting and simulations.

### 2022-12-05 background.tex
jaya

\acf{WCA} applications represent a category of novel, context-sensitive and highly-interactive \ac{AR} applications.
In this work we focus on a particular category --- denoted ``step-based'' \ac{WCA} --- that have as their goal the guiding of a user through a sequential task.
Examples of such applications are the LEGO and IKEA assistants~\cite{Chen2015LEGO,Chen2018application}, in which users are guided step-by-step through the process of assembling a LEGO model and an IKEA lamp, respectively.

Step-based \acp{WCA} operate analogously to how \ac{GPS} navigation assistants guide users, by seamlessly and continuously monitoring the progress of the user and autonomously providing relevant instructions and feedback.
The application follows the progress of the task in ``realtime'' by repeatedly sampling the state of the physical system, most commonly through video frames.
Whenever the assistant application detects that the user has correctly or incorrectly performed an instruction, it provides a new instruction to either advance the task or correct the detected mistake.
The application otherwise remains silent and out-of-the-way of the user; that is, samples which do not generate a new instruction (e.g. because they captured an intermediate or unfinished state, or simply noise) are silently discarded.
Herein lies one of key characteristics of these applications: the user only consciously interacts with the application whenever they finish an instruction, and thus these are the \emph{only} points in time at which they can notice changes in system responsiveness.

@@

In \textcite{olguinmunoz:impact2021} we studied the effects of system responsiveness on human behavior in step-based \ac{WCA} in a controlled experiment.
We employed a modified and instrumented version of the above-mentioned, step-based LEGO \ac{WCA}~\cite{Chen2015LEGO}.
Subjects interacted with the assistant while we altered the responsiveness in realtime and captured key application and task performance metrics.
Additionally, we also employed questionnaires to evaluate personality traits of the participants, and correlated these results with the task performance metrics.

@@

We recorded variables related to well-known individual differences, encompassing both the \acf{BFI}~\cite{oliver:bfi1999},and the \acf{ITQ}~\textcite{witmer1998measuring}.
Out of the individual-difference variables, the most salient effect on performance corresponded to \emph{neuroticism}, a \ac{BFI} trait linked to low tolerance for stress and high emotional reactivity, and which has previously been linked to higher \emph{delay discounting} rates~\cite{hirsh2008delay}.
Delay discounting is the tendency to devalue rewards for which one must wait; high rates, indicative of waiting intolerance, have been associated with negative social and academic outcomes.

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/new_model/mu_fits_exgaussian_slice0.png}
    \caption{%
        \( \mu \) parameter of \acs*{exGaussian} distributions fitted to execution times of the first four steps of segments of steps subject to the same \ac{TTF} in \textcite{olguinmunoz:impact2021}.
        Distributions were fitted using \ac{MLE}.
    }\label{fig:muexgaussian}
\end{figure}

>>>>

\acf{WCA} applications represent a category of novel, context-sensitive and highly-interactive \ac{AR} applications.
In this work we focus on a particular category --- ``step-based'' \ac{WCA} --- that have as their goal the guiding of a user through a sequential task.
Examples of such applications are the LEGO and IKEA assistants~\cite{Chen2015LEGO,Chen2018application}, in which users are guided step-by-step through the process of assembling a LEGO model and an IKEA lamp, respectively.

Step-based \acp{WCA} operate analogously to how \ac{GPS} navigation assistants guide users, by seamlessly and continuously monitoring the progress of the user and autonomously providing relevant instructions and feedback.
The application follows the progress of the task in ``realtime'' by repeatedly sampling the state of the physical system, most commonly through video frames.
Whenever the assistant detects that the user has correctly or incorrectly performed an instruction, it provides a new instruction to either advance the task or correct the detected mistake.
The application otherwise remains silent and out-of-the-way of the user; that is, samples which do not generate a new instruction (e.g.~because they captured an intermediate or unfinished state, or simply noise) are silently discarded.
Herein lies one of the key characteristics of these applications: the user only consciously interacts with the application whenever they finish an instruction, and thus these are the \emph{only} points in time at which they can notice changes in system responsiveness.

@@

In~\cite{olguinmunoz:impact2021} we studied the effects of system responsiveness on human behavior in step-based \ac{WCA} in a controlled experiment.
We employed a modified and instrumented version of the above-mentioned, step-based LEGO \ac{WCA}~\cite{Chen2015LEGO}.
\num{40} subjects interacted with the assistant, performing a \num{169}-step task while we altered the responsiveness in realtime and captured key application and task performance metrics.
Additionally, we also employed questionnaires to evaluate personality traits of the participants, and correlated these results with the task performance metrics.

@@

We recorded variables related to well-known individual differences, encompassing both the \acf{BFI}~\cite{oliver:bfi1999}, and the \acf{ITQ}~\cite{witmer1998measuring}.
Out of the individual-difference variables, the most salient effect on performance corresponded to \emph{neuroticism}, a \ac{BFI} trait linked to low tolerance for stress and high emotional reactivity, and which has previously been linked to higher \emph{delay discounting} rates~\cite{hirsh2008delay}.
Delay discounting is the tendency to devalue rewards for which one must wait; high rates, indicative of waiting intolerance, have been associated with negative social and academic outcomes.

@@

In this work, we will use a normalized scale to describe neuroticism, derived from the minimum and maximum obtainable values for this variable in the \ac{BFI}~\cite{oliver:bfi1999}.
\emph{Low} and \emph{high} neuroticism will refer to the \( [0.0, 0.5) \) and \( [0.5, 1.0] \) ranges respectively.

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figs/new_model/mu_fits_exgaussian_slice0.png}
    \caption{%
        \( \mu \) parameter of \acs*{exGaussian} distributions fitted to execution times of the first four steps of segments of steps subject to the same \ac{TTF} in~\cite{olguinmunoz:impact2021}.
        Distributions were fitted using \ac{MLE}.
    }\label{fig:muexgaussian}
\end{figure}

### 2022-12-05 conclusion.tex
jaya

The paper then explores the impact of such a realistic model on the application lifetime footprint of \ac{WCA} applications.
It shows that less realistic modeling approachesm which do not take into account higher-order effects on execution time distributions, can potentially lead to substantial mis-estimations of the application footprint.

Finally, the paper delves into the potential for optimization in \ac{WCA} systems using the previously discussed timing models.
It proposes a novel stochastic optimization framework for resource consumption-system responsiveness trade-offs in \ac{WCA}, which results in an adaptive sampling strategy. 
We have shown that this framework is applicable to a myriad of metrics in these applications, and showcased experimental results employing this framework for the minimization of number of samples processed per step and total energy consumption per step.
Our results show up to a \SI{50}{\percent} increase in performance with respect to state of the art when optimizing for number of samples, and up to a \SI{30}{\percent} improvement when optimizing for energy consumption, thus proving the value of such frameworks for the design of \ac{WCA} applications.

@@

\todo[inline]{Comment from Bobby: Reviewers might wonder why we stuck with the Rayleigh distribution.}

>>>>

The paper then explores the impact of such a realistic model on the application lifetime footprint of \ac{WCA} applications.
It shows that less realistic modeling approaches which do not take into account higher-order effects on execution time distributions, can potentially lead to substantial mis-estimations of the application footprint.

Finally, the paper delves into the potential for optimization in \ac{WCA} systems using the previously discussed timing models.
It proposes a novel stochastic optimization framework for resource consumption-system responsiveness trade-offs in \ac{WCA}, which results in an adaptive sampling strategy. 
We have shown that this framework is applicable to both the minimization of number of samples and total energy consumption per step by showcasing experimental results.
Our results show up to a \SI{50}{\percent} increase in performance with respect to state of the art when optimizing for number of samples, and up to a \SI{30}{\percent} improvement when optimizing for energy consumption, thus proving the value of such frameworks for the design of \ac{WCA} applications.

@@

\todo[inline]{Comment from Bobby: Reviewers might wonder why we stuck with the Rayleigh distribution.
Manuel: I think Vishnu already fixed this in the appendix.}

### 2022-12-05 implications_footprint.tex
jaya

In order to illustrate the consequences of using a less realistic model that does not take into account higher order effects, we introduce here a reference model to which we will compare our more realistic models.
This model represents a first-order approximation to empirical execution time modeling, and consist simply of an \ac{exGaussian} distribution fitted to all execution time samples collected for \textcite{olguinmunoz:impact2021}.
This distribution is then randomly sampled at runtime to obtain execution times for each step, without any adjustment to the current state of the system.

@@

We deploy configurations running \num{45}-step versions of the LEGO task described in \textcite{olguinmunoz:impact2021}.
The testbed configurations include setups with \numlist{1;4;7;10} clients.
Due to having limited time, each combination of testbed and timing model configuration is only repeated \num{10} times.
The results are presented in \cref{fig:testbed_lifetimes}.
Owing to the low number of samples, we opt here for the geometric rather than arithmetic mean to represent our results.
The results are nonetheless clear, and follow the same pattern as the previously discussed results under ideal, controlled circumstances.
With just a single client and mean \ac{RTT} of around \SI{400}{\milli\second}, all parameterizations of the realistic model achieved an average task duration \SIrange{6}{12}{\percent} shorter than the reference.
At \num{10} clients, the results mimic those at higher \acp{TTF} in the ideal controlled setup, with high neuroticism parameterizations reaching \textasciitilde\SI{8}{\percent} longer application lifetimes.

>>>>

In order to illustrate the consequences of using a less realistic model that does not take into account higher order effects, we introduce here a reference model to which we will compare our more realistic models.
This model represents a first-order approximation to empirical execution time modeling, and consist simply of an \ac{exGaussian} distribution fitted to all execution time samples collected for~\cite{olguinmunoz:impact2021}.
This distribution is then randomly sampled at runtime to obtain execution times for each step, without any adjustment to the current state of the system.

@@

We deploy configurations running \num{45}-step versions of the LEGO task described in~\cite{olguinmunoz:impact2021}.
The testbed configurations include setups with \numlist{1;4;7;10} clients.
Due to having limited time, each combination of testbed and timing model configuration is only repeated \num{10} times.
The results are presented in \cref{fig:testbed_lifetimes}.
Owing to the low number of samples, specifically to minimize the effects of potential outliers, we opt here for the geometric rather than arithmetic mean to represent our results.
The results are nonetheless clear, and follow the same pattern as the previously discussed results under ideal, controlled circumstances.
With just a single client and mean \ac{RTT} of around \SI{400}{\milli\second}, all parameterizations of the realistic model achieved an average task duration \SIrange{6}{12}{\percent} shorter than the reference.
At \num{10} clients, the results mimic those at higher \acp{TTF} in the ideal controlled setup, with high neuroticism parameterizations reaching \textasciitilde\SI{8}{\percent} longer application lifetimes.

### 2022-12-05 implications_optimization.tex
jaya

% In this optimization, the constant parts are excluded as it does not affect the process.
With a value much smaller than the execution time of the event, the \ac{RTT} of the final sample is included in this constant part, which is why $\mathcal{W}$ is computed without it.
Next, in \textcite{Moothedath2022Aperiodic}, the author retains the model, removes the constraint of periodicity, and finds the optimum aperiodic sampling instants $\{t_n,\,n=1,2,\dots\}$ that minimize the same energy penalty.

In this work, we use the modeling from~\cite{Moothedath2022Aperiodic} to find the optimum \emph{aperiodic} sampling interval.
However, instead of using their two-step approach to find the solution which includes a recursive solution followed by a bisection algorithm, we develop a novel, approximate, but easier solution to finding the set of the optimum aperiodic sampling intervals.
Furthermore, instead of directly optimizing for energy, we start by noting that an objective metric of a variety of optimization problems that are related to sampling, RTT, responsiveness or wait time present themselves 
% in a format similar to that of the energy, 
as a linear combination between $\mathbb{E}[\mathcal{S}]$ and $\mathbb{E}[\mathcal{W}]$ plus terms independent of the number of samples or wait time.
Let $\mathcal{E}$ correspond to such an objective metric.
Thus, 
\begin{alignat}{1}
    \Rightarrow\mathcal{E}&=\alpha\mathbb{E}[\mathcal{S}]+\beta\mathbb{E}[\mathcal{W}]+C,\;\label{eq:epsilon_terminal}
\end{alignat}

Above, $\alpha, \beta$ and $C$ are constants which are responsible for modifying the objective function from one metric to another.
For instance, in the modeling used by \textcite{Moothedath2021EnergyOptimal,Moothedath2022EnergyEfficient,Moothedath2022Aperiodic}, the chosen characterization results in a metric which is equal to the energy penalty.
Our solution which minimizes \cref{eq:epsilon_terminal} is provided in \cref{tn_approx_rayleigh}.
It assumes a $\mathcal{T}$ distributed according to a Rayleigh distribution with parameter $\sigma$.
The complete mathematical derivation that leads to this formula is detailed in \cref{appx1}.

\begin{alignat}{1}
r^*(t)&=\sqrt{\dfrac{\beta t}{2\alpha\sigma^2}}\nonumber\\
\Rightarrow t_n&=\Big(3\sigma\!\sqrt{\tfrac{\alpha}{2\beta}}\Big)^{\frac{2}{3}}n^{\frac{2}{3}}\label{tn_approx_rayleigh}
\end{alignat}

@@

However, an unconstrained optimization of the number of samples is trivial and meaningless as the solution points to a single sample at $t\!\rightarrow\!\infty$, which also takes the wait time to infinity.
Thus, we look at the constrained optimization of the expected number of samples with an upper bound $w_0$ for the expected wait.
That is, $\mathbb{E}[\mathcal{W}]\!\leq\!w_0$. We show in \cref{appx1} that the appropriate $\alpha$ and $\beta$ for this problem satisfy the condition
\begin{alignat}{1}\label{eq:optimization:sampling}
\frac{\alpha}{\beta}=\frac{2\sqrt{2}\,w_0^2}{(\mathlarger{\Gamma}(\tfrac{3}{4}))^2\,\sigma}\approx1.9\frac{w_0^2}{\sigma}
\end{alignat}
where, $\mathlarger{\Gamma}(x)$ is the Gamma function.

@@

In the following, we will show the effects of our optimization approach combined with our timing models compared to \textcite{Wang2019Towards}'s state-of-the-art approach.
For this we implement these sampling schemes in Python and run a number of simulations with them.
The first scheme uses our approach, \cref{eq:epsilon_terminal,eq:optimization:sampling}, to determine the optimum sampling instants and the values for \( \alpha \) and \( \beta \) at each step.
It includes an embedded timing model (without any distribution fitting) to provide updated estimates of the mean execution time \( \mu \) and \( \sigma \) at every step as well, allowing it to adapt to the state of the system.
We will refer to this scheme as the \emph{sample-count-optimized aperiodic} sampling scheme.

@@

It should be noted that although both schemes seem to tend towards \num{2.0} samples per step as \acp{RTT} increase, this is simply an artifact of our experimental setup.
As \acp{RTT} increase above the expected wait time, the probability of the sum of the first sampling interval and the \ac{RTT} being larger than the execution time of the step tends towards \num{1.0}.
This leads to these sampling schemes consistently sampling only twice each step: a first sample which is taken before the execution time of the step, and a second one \ac{RTT} seconds later, after the execution time has been reached.

@@

In the following, we thus take the general solution, \cref{eq:epsilon_terminal}, and find the appropriate \( \alpha \) and \( \beta \) to minimize the energy consumed per step, i.e.\ \( \mathcal{E}=E \).
We directly take the modelling from~\cite{Moothedath2022Aperiodic} with a necessary modification in the  assumption of one-way communication in all but the final sample.
With feedback given even to the discarded samples in our model and communication delay defined as the total delay in either direction, we have, 
\begin{alignat}{2}
    \mathrm{E}=&\;\mathcal{S}\tau_cP_c+(\mathcal{T}+\mathcal{W}+\tau_\mathrm{p}+\tau_\mathrm{c}-\mathcal{S}\tau_c)P_0\nonumber\\
    % &=s\tau_c(P_c-P_0^{(t)})+wP_0^{(t)}+(\tau+\tau_c+\tau_p) P_0^{(t)}+\tau_cP_c\\
    =&\;\tau_{\text{c}}(P_{\text{c}} -P_0)\mathcal{S}+\mathcal{W}P_0+(\mathcal{T}+\tau_{\text{p}} +\tau_{\text{c}}) P_0\nonumber\\
&\Rightarrow \alpha=\tau_{\text{c}}(P_{\text{c}} -P_0),\text{ and }\beta=P_0 \label{eq:optimization:energy}
\end{alignat}
\( \tau_\text{p} \) and \( \tau_\text{c} \) correspond to the processing and two-way communication delay for each sample.
\( P_\text{c} \) and \( P_0 \) correspond to the communication and idle power, respectively, of the \ac{WCA} client device.

We proceed to repeat the experiment detailed in \cref{ssec:optimization:samples}, replacing the sample-count-optimized sampling scheme with a new implementation instead minimizing energy, using \cref{eq:optimization:energy} for the calculation of \( \alpha \) and \( \beta \).
Once again, we embed a timing model into the sampling scheme to provide updated estimates of the mean execution time \( \mu \) at each step.
We refer to this scheme as the \emph{energy-optimized aperiodic} sampling scheme.
For the power constants, we reuse the values estimated by the authors in \textcite{Moothedath2022EnergyEfficient}, \( P_\text{c} = 45\,\si{\milli\watt} \) and \( P_\text{0} = 15\,\si{\milli\watt} \).
On the other hand, for the timing variables we define a constant processing delay \( \tau_\text{p} = 250\,\si{\milli\second} \) across all configurations and repetitions of the experiments; given then a constant \ac{RTT} for the task, we set \( \tau_\text{c} = \text{\ac{RTT}} - \tau_\text{p} \).

@@

The results of these experiments are presented in \cref{fig:optimization:energy}, and they clearly illustrate the advantages of the integration of our timing models with an adaptive, energy-optimized sampling scheme when compared to unoptimized and state-of-the-art sampling schemes.
Our approach is consistently consumes \SI{20}{\percent} less energy than \textcite{Wang2019Towards}'s state-of-the-art, and is up to \SI{40}{\percent} more efficient than the greedy scheme.
Furthermore, once again the behavior of our approach is more consistent and reliable than the competition, exhibiting a flat curve of energy consumption much akin to that of the offline optimum in behavior.

>>>>

% In this optimization, the constant parts are excluded as it does not affect the process.
With a value much smaller than the execution time of the event, the \ac{RTT} of the final sample is included in this constant part, which is why $\mathcal{W}$ is computed without it.
Next, in~\cite{Moothedath2022Aperiodic}, the author retains the model, removes the constraint of periodicity, and finds the optimum aperiodic sampling instants $\{t_n,\,n=1,2,\dots\}$ that minimize the same energy penalty.

In this work, we use the modeling from~\cite{Moothedath2022Aperiodic} to find the optimum \emph{aperiodic} sampling interval.
However, instead of using their two-step approach to find the solution which includes a recursive solution followed by a bisection algorithm, we develop a novel, approximate, but easier solution to finding the set of the optimum aperiodic sampling intervals.
Furthermore, instead of directly optimizing for energy, we start by noting that, in general, any objective metric in these applications which relates to sampling and the responsiveness of the system will present itself
% that an objective metric of a variety of optimization problems that are related to sampling, RTT, responsiveness or wait time present themselves 
% % in a format similar to that of the energy, 
as a linear combination between $\mathbb{E}[\mathcal{S}]$ and $\mathbb{E}[\mathcal{W}]$ plus terms independent of the number of samples or wait time.
Let $\mathcal{E}$ correspond to such an objective metric.
Thus, 
\begin{alignat}{1}
    \Rightarrow\mathcal{E}&=\alpha\mathbb{E}[\mathcal{S}]+\beta\mathbb{E}[\mathcal{W}]+C\;\label{eq:epsilon_terminal}
\end{alignat}

\todo[inline]{%
    Comment by Jaya:
    The presentation of the optimization for energy metric in (5) and later in Section 6.2 presenting the optimization for minimizing number of samples subject to weight penalty is done in a non-trivial way. A natural way to present it is to state the optimization problem in 6.2 first and state that the function in (5) is a Legrangian function of this optimization with alpha/beta being the Lagrangian parameter. People are well aware of the Lagrangian formulation, so at least mention that (5) is a Lagrangian function and alpha/beta is a Lagrangian parameter.

    Manuel: I've added a mention to the Lagrangian function below. Do we need anything else?
}

This corresponds to a Lagrangian function of this optimization, with \ensuremath{\frac{\alpha}{\beta}} representing the Lagrangian parameter.
$\alpha, \beta$ and $C$ are constants which are responsible for modifying the objective function from one metric to another.
For instance, in the modeling used by~\cite{Moothedath2021EnergyOptimal,Moothedath2022EnergyEfficient,Moothedath2022Aperiodic}, the chosen characterization results in a metric which is equal to the energy penalty.
Our solution which minimizes \cref{eq:epsilon_terminal} is provided in \cref{tn_approx_rayleigh}.
It assumes a $\mathcal{T}$ distributed according to a Rayleigh distribution with parameter $\sigma$.
The complete mathematical derivation that leads to this formula is detailed in \cref{appx1}.

\begin{alignat}{1}
% r^*(t)&=\sqrt{\dfrac{\beta t}{2\alpha\sigma^2}}\nonumber\\
% \Rightarrow
t_n&=\Big(3\sigma\!\sqrt{\tfrac{\alpha}{2\beta}}\Big)^{\frac{2}{3}}n^{\frac{2}{3}}\label{tn_approx_rayleigh}
\end{alignat}

@@

However, an unconstrained optimization of the number of samples is trivial and meaningless as the solution points to a single sample at $t\!\rightarrow\!\infty$, which also takes the wait time to infinity.
Thus, we look at the constrained optimization of the expected number of samples with an upper bound $w_0$ for the expected wait.
That is, $\mathbb{E}[\mathcal{W}]\!\leq\!w_0$.
We show in \cref{appx1} that we can find appropriate $\alpha$ and $\beta$ for this problem by satisfying the condition
\begin{alignat}{1}\label{eq:optimization:sampling}
\frac{\alpha}{\beta}=\frac{2\sqrt{2}\,w_0^2}{(\mathlarger{\Gamma}(\tfrac{3}{4}))^2\,\sigma}\approx1.9\frac{w_0^2}{\sigma}
\end{alignat}
where $\mathlarger{\Gamma}(x)$ is the Gamma function.

@@

In the following, we will show the effects of our optimization approach combined with our timing models compared to the state-of-the-art approach~\cite{Wang2019Towards}.
For this we implement these sampling schemes in Python and run a number of simulations with them.
The first scheme uses our approach, \cref{eq:epsilon_terminal,eq:optimization:sampling}, to determine the optimum sampling instants and the values for \( \alpha \) and \( \beta \) at each step.
It includes an embedded timing model (without any distribution fitting) to provide updated estimates of the mean execution time \( \mu \) and \( \sigma \) at every step as well, allowing it to adapt to the state of the system.
We will refer to this scheme as the \emph{sample-count-optimized aperiodic} sampling scheme.

@@

It should be noted that although both schemes seem to tend towards two samples per step as \acp{RTT} increase, this is simply an artifact of our experimental setup.
As \acp{RTT} increase above the expected wait time, the probability of the sum of the first sampling interval and the \ac{RTT} being larger than the execution time of the step tends towards \num{1.0}.
This leads to these sampling schemes consistently sampling only twice each step: a first sample which is taken before the execution time of the step, and a second one \ac{RTT} seconds later, after the execution time has been reached.

@@

In the following, we thus take the general solution, \cref{eq:epsilon_terminal}, and find the appropriate \( \alpha \) and \( \beta \) to minimize the energy consumed per step, i.e.\ \( \mathcal{E}=E \).
We directly take the modeling from~\cite{Moothedath2022Aperiodic} with a necessary modification in the  assumption of one-way communication in all but the final sample.
With feedback given even to the discarded samples in our model and communication delay defined as the total delay in either direction, we have, 
\begin{alignat}{2}
    \mathrm{E}=&\;\mathcal{S}\tau_cP_c+(\mathcal{T}+\mathcal{W}+\tau_\mathrm{p}+\tau_\mathrm{c}-\mathcal{S}\tau_c)P_0\nonumber\\
    % &=s\tau_c(P_c-P_0^{(t)})+wP_0^{(t)}+(\tau+\tau_c+\tau_p) P_0^{(t)}+\tau_cP_c\\
    =&\;\tau_{\text{c}}(P_{\text{c}} -P_0)\mathcal{S}+\mathcal{W}P_0+(\mathcal{T}+\tau_{\text{p}} +\tau_{\text{c}}) P_0\nonumber\\
&\Rightarrow \alpha=\tau_{\text{c}}(P_{\text{c}} -P_0),\text{ and }\beta=P_0 \label{eq:optimization:energy}
\end{alignat}
\( \tau_\text{p} \) and \( \tau_\text{c} \) correspond to the processing and two-way communication delay for each sample.
\( P_\text{c} \) and \( P_0 \) correspond to the communication and idle power, respectively, of the \ac{WCA} client device.

We proceed to repeat the experiment detailed in \cref{ssec:optimization:samples}, replacing the sample-count-optimized sampling scheme with a new implementation instead minimizing energy, using \cref{eq:optimization:energy} for the calculation of \( \alpha \) and \( \beta \).
Once again, we embed a timing model into the sampling scheme to provide updated estimates of the mean execution time \( \mu \) at each step.
We refer to this scheme as the \emph{energy-optimized aperiodic} sampling scheme.
For the power constants, we reuse the values estimated by the authors in~\cite{Moothedath2022EnergyEfficient}, \( P_\text{c} = 45\,\si{\milli\watt} \) and \( P_\text{0} = 15\,\si{\milli\watt} \).
On the other hand, for the timing variables we define a constant processing delay \( \tau_\text{p} = 250\,\si{\milli\second} \) across all configurations and repetitions of the experiments; given then a constant \ac{RTT} for the task, we set \( \tau_\text{c} = \text{\ac{RTT}} - \tau_\text{p} \).

@@

The results of these experiments are presented in \cref{fig:optimization:energy}, and they clearly illustrate the advantages of the integration of our timing models with an adaptive, energy-optimized sampling scheme when compared to unoptimized and state-of-the-art sampling schemes.
Our approach is consistently consumes \SI{20}{\percent} less energy than \textcite{Wang2019Towards}'s state-of-the-art, and is up to \SI{40}{\percent} more energy efficient than the greedy scheme.
Furthermore, once again the behavior of our approach is more consistent and reliable than the competition, exhibiting a flat curve of energy consumption much akin to that of the offline optimum in behavior.
